### 2015-06-25 Paper/Figures/FigTexture1D.tex
Paper: added accuracy diagrams, combined glsl listings

\documentclass{article}
\usepackage{tikz}
\usepackage{tkz-euclide}
\definecolor{darkgreen}{RGB}{0,192,0}
\title{Cartesian closed categories and the price of eggs}
\author{Jane Doe}
\date{September 1994}
\begin{document}

	\begin{figure}
		\begin{tikzpicture}[x=8cm,y=4cm]
			\tkzInit[xmax=2,ymax=1,xmin=0,ymin=0]
			% border
			\draw (-0.1,-0.2) rectangle (2,1.2);

			\foreach \x in {0,...,4}
				\draw (\x * 0.1 - 0.05,0.45) rectangle (\x * 0.1 + 0.05, 0.55);

		\end{tikzpicture}
		\caption{TODO: this}		
	\end{figure}

\end{document}